---
title: "NOAA Storm Data Analysis"
author: "krrish16"
date: "24 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database (start in the year 1950 and end in November 2011). This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

## Data Processing
```{r}
# Loading data
data <- read.csv("./StormData.csv.bz2")
dim(data)
head(data)

# Checking for Missing values
sapply(data , function(x) sum(is.na(x)))
```
Alphabetical characters are used to signify magnitude of property damage and crop damage that include "K" for thousands, "M" for millions, and "B" for billions.So we have to convert them into numericals
```{r}
# Converting property damage and crop damage
unique(data$PROPDMGEXP)

data$PROPEXP[data$PROPDMGEXP == "K"] <- 1000
data$PROPEXP[data$PROPDMGEXP == "M"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == ""]  <- 1
data$PROPEXP[data$PROPDMGEXP == "B"] <- 1e+09
data$PROPEXP[data$PROPDMGEXP == "m"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == "0"] <- 1
data$PROPEXP[data$PROPDMGEXP == "5"] <- 1e+05
data$PROPEXP[data$PROPDMGEXP == "6"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == "4"] <- 10000
data$PROPEXP[data$PROPDMGEXP == "2"] <- 100
data$PROPEXP[data$PROPDMGEXP == "3"] <- 1000
data$PROPEXP[data$PROPDMGEXP == "h"] <- 100
data$PROPEXP[data$PROPDMGEXP == "7"] <- 1e+07
data$PROPEXP[data$PROPDMGEXP == "H"] <- 100
data$PROPEXP[data$PROPDMGEXP == "1"] <- 10
data$PROPEXP[data$PROPDMGEXP == "8"] <- 1e+08
data$PROPEXP[data$PROPDMGEXP == "+"] <- 0
data$PROPEXP[data$PROPDMGEXP == "-"] <- 0
data$PROPEXP[data$PROPDMGEXP == "?"] <- 0
data$PROPDMGVAL <- data$PROPDMG * data$PROPEXP
head(data$PROPDMGVAL)

# Similarly converting crop damage
unique(data$CROPDMGEXP)

data$CROPEXP[data$CROPDMGEXP == "M"] <- 1e+06
data$CROPEXP[data$CROPDMGEXP == "K"] <- 1000
data$CROPEXP[data$CROPDMGEXP == "m"] <- 1e+06
data$CROPEXP[data$CROPDMGEXP == "B"] <- 1e+09
data$CROPEXP[data$CROPDMGEXP == "0"] <- 1
data$CROPEXP[data$CROPDMGEXP == "k"] <- 1000
data$CROPEXP[data$CROPDMGEXP == "2"] <- 100
data$CROPEXP[data$CROPDMGEXP == ""] <- 1
data$CROPEXP[data$CROPDMGEXP == "?"] <- 0
data$CROPDMGVAL <- data$CROPDMG * data$CROPEXP
head(data$CROPDMGVAL)
```

## Answering the given questions

1.Across the United States, which types of events (as indicated in the "EVTYPE") are most harmful with respect to population health?

### Aggregating and Sorting
```{r}

fatalities <- aggregate(FATALITIES ~ EVTYPE, data=data, FUN = sum)
fatalities <- fatalities[order(fatalities$FATALITIES, decreasing=TRUE),]
injuries <- aggregate(INJURIES ~ EVTYPE, data=data, FUN = sum)
injuries <- injuries[order(injuries$INJURIES, decreasing=TRUE),]
head(fatalities)
head(injuries)
```
### Plot 1 
```{r}
par(mfrow=c(1,2))
p1 <- barplot (height = fatalities$FATALITIES[1:10], names.arg = fatalities$EVTYPE[1:10], las = 2, cex.names= 0.6, col = "light blue" , main = "Fatalities Count vs Event Type")

p2 <- barplot (height = injuries$INJURIES[1:10], names.arg = injuries$EVTYPE[1:10], las = 2, cex.names= 0.6, col = "light blue" , main = "Injuries Count vs Event Type")
```

2.Across the United States, which types of events have the greatest economic consequences?

### Aggregating and Sorting
```{r}
property <- aggregate(PROPDMGVAL ~ EVTYPE, data, FUN = sum)
property <- property[order(property$PROPDMGVAL, decreasing=TRUE),]
crop <- aggregate(CROPDMGVAL ~ EVTYPE, data, FUN = sum)
crop <- crop[order(crop$CROPDMGVAL, decreasing=TRUE),]
head(property)
head(crop)
```

### Plot 2
```{r}
par(mfrow=c(1,2))
barplot (height = property$PROPDMGVAL[1:10]/(10^9), names.arg = property$EVTYPE[1:10], las = 2, cex.names= 0.6,col = "blue" , main = "Property Damage vs Event Type")

barplot (height = crop$CROPDMGVAL[1:10]/(10^9), names.arg = crop$EVTYPE[1:10], las = 2, cex.names= 0.6,col = "blue" , main = "Crop Damage vs Event Type")
```


## Results
From Plot 1 , 
Tornados are the most harmful events that caused maximum number of fatalities and injuries followed by Excessive Heat and Thunderstorm wind in case of fatalities and injuries respectively. 

From Plot 2 ,
Floods are the events that have the greatest economic consequences causing the maximum property damage while Drought caused the maximum crop damage followed by Hurricanes and Floods for property damage and crop damage respectively