---
title: "Week 2 Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data

```{r}
library(ggplot2)
data <- read.csv("./activity.csv")
head(data)
```

## Missing Values

```{r}
sapply(data, function(x) sum(is.na(x)))
```

## Removing NA's

```{r}
na_rows <- which(is.na(data$steps))
re_na_data <- data[-na_rows,]
```

## Number of Steps per day

```{r}
stepsperday <- aggregate(steps ~ date,data = data , FUN = sum)
head(stepsperday)
```

## Histogram

```{r}
ggplot(stepsperday, aes(x = steps))+geom_histogram(fill = "blue", binwidth=1000)+labs(title = "Steps Per Day", x = "Steps", y = "Frequency")
```

## Mean and Median of the total number of steps taken per day

```{r}
mean_steps <- mean(stepsperday$steps)
median_steps <- median(stepsperday$steps)
print(paste("mean : ", round(mean_steps,2)))
print(paste("median : ", round(median_steps,2)))
```

## Average daily activity pattern

```{r}
interval_steps <- aggregate(steps ~ interval,data = re_na_data , FUN =mean)
ggplot(interval_steps, aes(x = interval, y=steps))+geom_line(color="blue")+labs(title = "Steps by Interval", x = "interval", y = "steps")
```

## 5 min interval with max no. of steps

```{r}
max_interval <- interval_steps[which.max(interval_steps$steps),]
max_interval
```

## Number of missing values

```{r}
sapply(data, function(x) sum(is.na(x)))
```

## Filling NA's with mean of steps and creating a new Data set

```{r}
new_data <- data
Val <- tapply(data$steps, data$interval, mean, na.rm=TRUE, simplify=TRUE)
new_data$steps[na_rows] <- Val[as.character(new_data$interval[na_rows])]
```

## New Data Histogram

```{r}
new_stepsperday <- aggregate(steps ~ date,data = new_data , FUN = sum)
head(new_stepsperday)
ggplot(new_stepsperday, aes(x = steps))+geom_histogram(fill = "blue", binwidth=1000)+labs(title = "Steps Per Day", x = "Steps", y = "Frequency")
```

## Mean and Median of the total number of steps taken per day for new data

```{r}
new_mean_steps <- mean(new_stepsperday$steps)
new_median_steps <- median(new_stepsperday$steps)
print(paste("mean : ", round(new_mean_steps,2)))
print(paste("median : ", round(new_median_steps,2)))
```
### Here the mean value for both cases are equal but in case of median there's a slight variation

## Differences in activity patterns
```{r}
new_data$date <- as.Date(as.character(data$date))
new_data$day <- weekdays(new_data$date)
#?ifesle
new_data$is_weekend <- ifelse(new_data$day=="Saturday" | new_data$day=="Sunday","yes", "no")
weekends <- aggregate(steps ~ is_weekend + interval,data = new_data,FUN = mean)
ggplot(weekends, aes(x = interval, y=steps, color= is_weekend)) + geom_line() + facet_grid(is_weekend ~ .) + labs(title = "Mean of Steps by Interval", x = "interval", y = "steps")
```

### It seems to be that there are more number of steps during weekends